\pdfoutput=1\relax
\documentclass[reqno]{amsart}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\input{Preamble.tex}
\usepackage[margin=1.5in]{geometry}

\newtoggle{draft}
% NB: toggle this to hide comments
%\togglefalse{draft}
\toggletrue{draft}

\iftoggle{draft}{
%\usepackage{showkeys}
\newcommand{\atodo}[1]{\todo[color=blue!40]{#1}}
\newcommand{\ttodo}[1]{\todo[color=green!40]{#1}}
\newcommand{\tcomment}[1]{\todo[inline,color=green!40]{#1}}
\newcommand{\NB}[1]{\todo[color=gray!40]{#1}}
\newcommand{\TODO}[1]{\todo[color=red]{#1}}
}{ % else
\newcommand{\NB}[1]{}
\newcommand{\TODO}[1]{}
\renewcommand{\todo}[1]{}
\renewcommand{\todo}[1]{}
}

\title{Exercise sheet 4}

%% \author{Robert Burklund}
%% \address{Department of Mathematical Sciences, University of Copenhagen, Denmark}
%% \email{rb@math.ku.dk}

\begin{document}
\maketitle


\section*{\bf Background}

The first $13$ stable homotopy groups of the $2$-complete sphere $\Ss_{2}$ can be computed using only the Serre spectral sequence. 
These groups (and the ring structure) are given in the table below:

\begin{center}
  \begin{tabular}{|c|c|c|c|}\hline
    stem & group & generator & relations \\\hline\hline
    0 & $\Z_2$ & $1$ & \\\hline
    1 & $\Z/2$ & $\eta$ & \\\hline
    2 & $\Z/2$ & $\eta^2$ & \\\hline
    3 & $\Z/8$ & $\nu$ & $\eta^3 = 4\nu$ \\\hline
    4 & $0$ & & \\\hline
    5 & $0$ & & \\\hline
    6 & $\Z/2$ & $\nu^2$ & \\\hline
    7 & $\Z/16$ & $\sigma$ & \\\hline
    8 & $\Z/2 \oplus \Z/2$ & $\eta\sigma$, $\epsilon$ & \\\hline
    9 & $\Z/2 \oplus \Z/2 \oplus \Z/2$ & $\eta^2\sigma$, $\eta\epsilon$, $\mu_9$ & $\nu^3 = \eta^2\sigma + \eta\epsilon$ \\\hline
    10 & $\Z/2$ & $\eta\mu_9$ & \\\hline
    11 & $\Z/8$ & $\zeta$ & $\eta^2\mu_9 = 4\zeta$ \\\hline
    12 & $0$ & & \\\hline
    13 & $0$& & \\\hline
  \end{tabular}
\end{center}


\section{\bf Exercise 1: The homotopy of $\KO$ (hfp approach)}

$\mathrm{KO}$ is an $\E_\infty$-algebra in $\Sp$ whose associated cohomology theory
classifies stable vector bundles.
Similarly, $\mathrm{KU}$ is an $\E_\infty$-algebra in $\Sp$ whose associated cohomology theory
classifies stable complex vector bundles.
Complexification of vector bundles induces a map of $\E_{\infty}$-algebras
$\KO \to \KU,$
furthermore complex conjugation provides a $C_2$-action on $\KU$ such that
$\KO \cong \KU^{hC_2}$.
The $C_2$ action on $\KU$ by complex conjugation
sends the generator $\beta$ in $\pi_2\KU$ to $-\beta$.

\begin{itemize}
\item[(bonus)] Using your understanding of complex vector bundles prove that
  complex conjugatin acts by $-1$ on $\pi_2\KU$.
\item[(a)] Determine the $E_2$-page of the $C_2$-homotopy fixed point sseq which has signature
  \[ E_2^{s,t} \Rightarrow \pi_{t-s}(\KU^{hC_2}). \]
\end{itemize}

In order to proceed we will need one more piece of information:
The image of the class $\eta \in \pi_1\Ss$ in $\pi_1\KO$ is non-zero and
moreover this class is detected in the $C_2$-hfp sseq in $E_\infty^{1,2}$.
With this information in hand we will now investigate the $C_2$-homotopy fixed point sseq.

\begin{itemize}
\item[(b)] Determine the differentials in the $C_2$-hfp sseq for $\KU^{hC_2}$.
\item[(c)] Determine the ring of homotopy groups $\pi_*\KO$.
\item[(d)] Determine the bigraded homotopy ring of the filtered ring spectrum
  $(\tau_{\geq \bullet} \KU)^{hC_2}$ underlying the $C_2$-hfp sseq.
\end{itemize}

\section{\bf Exercise 2: Slicing sseqs}

\begin{itemize}
\item[(a)] Determine the homotopy groups of $\ku^{hC_2}$ where the action is by complex conjugation (recall that $\ku$ is the connective cover of $\KU$). Observe that $\ku^{hC_2}$ is not $\ko$.
\end{itemize}

In the remainder of this exercise we will explore the idea of ``slicing'' spectral sequences.
Let $\left( {}^AE_r^{s,t}, d_r \right)$ and $\left( {}^BE_r^{s,t}, d_r \right)$
be a pair of sseqs and suppose we have a  map
\[ \left( {}^AE_r^{s,t}, d_r \right) \to \left( {}^BE_r^{s,t}, d_r \right) \]
between them.

\begin{itemize}
\item[(b)] Let $(m,c)$ be.
  Suppose that the map of $E_2$-pages is an isomorphism
  for $s \geq m(t-s) + c$ and
  ${}^A E_2^{s,t} = 0$ for $s < m(t-s) + c$.
  For which $(m,c)$ can the differentials in the $A$-sseq be deduced from the
  differentials in the $B$-sseq.
\item[(c)] Under the hypotheses of (b)
  with $(m=-1,c=0)$ give a formula for ${}^AE_\infty^{s,t}$ in terms of the $B$-sseq.
\item[(d)] Let $X$ be a spectrum with a $G$-action, taking connective covers we
  obtain a $G$-action on $\tau_{\geq 0}X$ and a $G$-equivariant map $\tau_{\geq 0}X \to X$.
  Show that the associated map of $G$-homotopy fixed point sseqs
  is an instance of slicing.
\end{itemize}

Now we'll think about the dual situation.

\begin{itemize}
\item[(e)] Let $(m,c)$ be.
  Suppose that the map of $E_2$-pages is an isomorphism
  for $s \leq m(t-s) + c$ and
  ${}^B E_2^{s,t} = 0$ for $s > m(t-s) + c$.
  For which $(m,c)$ can the differentials in the $B$-sseq be deduced from the
  differentials in the $A$-sseq.
\item[(f)] Under the hypotheses of (d)
  with $(m=-1,c=0)$ give a formula for ${}^AE_\infty^{s,t}$ in terms of the $B$-sseq.  
\item[(g)] Let $X_\bullet$ be a spectrum with a skeletal filtration.
  Let $(X_k)_{\bullet}$ be the $k$-skeleton of $X$ with its restricted skeletal filtation.
  Show that the map of skeletal sseqs associated to $(X_k)_\bullet \to X_\bullet$
  is an instance of slicing.      
\end{itemize}
  
Think about some scenarios where a sseq might be sliced multiple times.

\section{\bf Exercise 3: Bockstein sseqs}

%% special properties of bockstein sseqs
A bockstein sseq is a sseq whose $E_2$-page is of the form
$E_2^{s,t} \cong E_2^{0,t} \otimes \Z[b]$
where the $s$-degree of $b$ is positive
and the diferentials are all $b$-linear.

In this exercise you will investigate the special features of bockstein sseqs.

\begin{itemize}
\item[(a)] Show that a bockstein sseq is uniquely determine by the differentials off the zero line (i.e. $s$-degree 0).
\item[(b)] Show that $b$ can be formally inverted on the $E_2$-page and that there is a correspond localized bockstein sseq.
\item[(c)] Show that the localized bockstein sseq uniquely determines the bockstein sseq.
\item[(d)] Using (c) show that it is possible to set $b=1$ in the localized bockstein sseq and to obtain singly-graded sseq (only the $t$-degree will remain) which continues to contain the full information of the bockstein sseq.
\end{itemize}

It is useful to compare part (d) with the discussion of the $\tau$-bockstein sseq on a filtered spectrum $X_\bullet$ from class.

\section{\bf Exercise 4: Synthetic stems}

In class we determined the first 13 synthetic stems.
These groups are given in the table below:

%% The first $13$ stable homotopy groups of the $2$-complete sphere $\Ss_{2}$ can be computed using only the Serre spectral sequence. 

Let $\pi_{0*}$ denote the graded ring
\[ \pi_{0*} \coloneqq \Z_2[\wt{2}, \tau]/ \wt{2}\tau = 2. \]
where $|\tau| = (0,-1)$ and $|\wt{2}| = (1,0)$.

\begin{center}
  \begin{tabular}{|c|c|c|c|}\hline
    stem & group & generator & degrees \\\hline\hline
    0 & $\pi_{0*}$ & $1$ & \\\hline
    1 & $\pi_{0*}/\wt{2}$ & $\eta$ & $|\eta| = (1,1)$ \\\hline
    2 & $\pi_{0*}/\wt{2}$ & $\eta^2$ & $|\eta^2| = (2,2)$ \\\hline
    3 & $\pi_{0*}/\wt{2}^3$ & $\nu$ & $|\nu| = (3,1)$ \\\hline
    4 & $0$ & & \\\hline
    5 & $0$ & & \\\hline
    6 & $\pi_{0*}/\wt{2}$ & $\nu^2$ & $|\nu^2| = (6,2)$ \\\hline
    7 & $\pi_{0*}/\wt{2}^4$ & $\sigma$ & $|\sigma| = (7,1)$ \\\hline
    8 & $\pi_{0*}/\wt{2} \oplus \pi_{0*}/\wt{2}$ & $\eta\sigma$, $\epsilon$ & $|\eta\sigma| = (8,2)$, $|\epsilon| = (8,3)$\\\hline
    9 & $\pi_{0*}/\wt{2} \oplus \pi_{0*}/\wt{2} \oplus \pi_{0*}/\wt{2}$ & $\eta^2\sigma$, $\eta\epsilon$, $\mu_9$ & $|\eta^2\sigma| = (9,3)$, $|\eta\epsilon| = (8,4)$, $|\mu_9| = (9,5)$ \\\hline
    10 & $\pi_{0*}/\wt{2}$ & $\eta\mu_9$ & $|\eta\mu_9| = (10,6)$ \\\hline
    11 & $\pi_{0*}/\wt{2}^3$ & $\zeta$ & $|\zeta| = (11, 5)$ \\\hline
    12 & $0$ & & \\\hline
    13 & $0$& & \\\hline
  \end{tabular}
\end{center}

The ext ring $\Ext_{\A}^{**}(\F_2,\F_2) \cong \pi_{**}C\tau$ has been determined in a large range through computer calculations.
See \href{http://www.rrb.wayne.edu/papers/index.html}{Bob Bruner's webpage} and particularly ``40. The cohomology of the mod 2 Steenrod Algebra'' for this data.

\begin{itemize}
\item[(a)] Using the Ext ring data and the $\tau$-bockstein sseq
  confirm (or correct) the entries in the table above.\footnote{Ignore convergence issues.}
\item[(b)] Compute the ring structure in this range.
\item[(c)] Tensoring with $C\tau$ gives a symmetric monoidal functor to an algebraic category where it should be easy to compute products and Toda brackets (though we have yet to introduce techniques to do so). Think through the details of why functors send Toda brackets to Toda brackets. Now re-examine the Toda brackets computed in Exercise sheet 2 and see how many of them you can reduce to an algebraic question.
\end{itemize}

\section{\bf Exercise 5: A bracket in $\KO$}

In this exercise you will recover an instance of Moss' theorem on
the interaction between Toda brackets and sseqs.

Prove that in $\pi_*\KO$ we have
\[ \langle 2, \eta, \eta^2 \rangle = \alpha + 2\blacksquare \]
where $\alpha$ is a generator of $\pi_4\KO$.

For this part it will be useful to think about the filtered $\E_\infty$-algebra
$(\tau_{\geq \bullet} \KU)^{hC_2}$ from Ex.1 and first evaluate a lift of this bracket to bigraded homotopy groups.
It will be useful to think about the ways in which powers of $\tau$ can be added in to brackets and how this intercts with the $\tau$-bockstein sequences.



\section{\bf Exercise 6: The Wood cofiber sequence (ver 2)}

$\mathrm{KO}$ is an $\E_\infty$-algebra in $\Sp$ whose associated cohomology theory
classifies stable vector bundles.
Complexification gives a map of $\E_\infty$-algebras
\[ \KO \to \KU. \]
In this exercise you will prove that $\cof(\eta) \otimes \KO \cong \KU$.\footnote{The same question appeared on the previous exercise sheet, but we will follow a different approach this time.}

\begin{itemize}
\item[(a)] Recall that the homotopy ring of $\KU$ is given by
  $\pi_*\KU \cong \Z[\beta^\pm]$ where $|\beta|=2$ and
  the homotopy ring of $\KO$ is given by
  \[ \pi_*\KO \cong \Z[\eta, \alpha, \gamma^\pm]/(2\eta, \eta^3, \eta\alpha, \alpha^2=4\gamma). \]
  Determine what the complexification map does on homotopy rings.
\item[(b)] Use the fact that $\nu$ is zero in $\pi_*\KO$ to give
  $\cof(\eta) \otimes \KO$ a unital multiplication. How many unital multiplications are there?
\item[(c)] Use the fact that $\eta$ is zero in $\pi_*\KU$ to show that there is a unital multiplication for which there is a multiplicative map
  $\cof(\eta) \otimes \KO \to \KU$.
\item[(d)] Compute the homotopy groups of $\cof(\eta) \otimes \KO$
  (use the bracket from the previous exercise).
\item[(e)] What is the square of the generator in degree $2$ of
  $\pi_*(\cof(\eta) \otimes \KO)$?
\item[(f)] Conclude that $\cof(\eta) \otimes \KO \cong \KU$.
\end{itemize}




\bibliographystyle{alpha}
\bibliography{bibliography}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% eval: (visual-line-mode t)
%%% eval: (auto-fill-mode 0)
%%% End:
