\pdfoutput=1\relax
\documentclass[reqno]{amsart}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\input{Preamble.tex}
\usepackage[margin=1.5in]{geometry}

\newtoggle{draft}
% NB: toggle this to hide comments
%\togglefalse{draft}
\toggletrue{draft}

\iftoggle{draft}{
%\usepackage{showkeys}
\newcommand{\atodo}[1]{\todo[color=blue!40]{#1}}
\newcommand{\ttodo}[1]{\todo[color=green!40]{#1}}
\newcommand{\tcomment}[1]{\todo[inline,color=green!40]{#1}}
\newcommand{\NB}[1]{\todo[color=gray!40]{#1}}
\newcommand{\TODO}[1]{\todo[color=red]{#1}}
}{ % else
\newcommand{\NB}[1]{}
\newcommand{\TODO}[1]{}
\renewcommand{\todo}[1]{}
\renewcommand{\todo}[1]{}
}

\title{Exercise sheet 2}

%% \author{Robert Burklund}
%% \address{Department of Mathematical Sciences, University of Copenhagen, Denmark}
%% \email{rb@math.ku.dk}

\begin{document}
\maketitle

\section*{\bf Exercise 1: Advanced shuffling}

In lecture we discussed the shuffling formulas
\begin{itemize}
\item $\langle a,b,cd \rangle \subseteq \langle a,bc,d \rangle$,
\item $\langle a,b,c \rangle d = -a \langle b,c,d \rangle$ and
\item $\langle a,b,c \rangle + \langle b,c,a \rangle + \langle c,a,b \rangle = 0$
\end{itemize}
and the situations in which they apply.
In this exercise you will prove two additional shuffling formulas.

Let $\CC$ be a stable category.
Let
\[ U \xrightarrow{a} V \xrightarrow{b} W \xrightarrow{c} X \xrightarrow{d} Y \]
be a sequence of four composable maps in $\CC$ and
suppose that all $ab$ and $bc$ are nullhomotopic.

\begin{enumerate}
\item[(a)] Show that
  \[ \langle a,b,c \rangle d \subseteq \langle a,b,cd \rangle. \]
\item[(b)] Identify which nullhomotopies one should use to obtain a equality.
\end{enumerate}

Let
\[ U \xrightarrow{a} V \xrightarrow{b} W \xrightarrow{c} X \xrightarrow{d} Y \xrightarrow{e} Z \]
be a sequence of five composable maps in $\CC$ and
let $e_2 : bc \leftrightarrow 0$ and $e_3 : cd \leftrightarrow 0 $ be nullhomotopies.
Suppose further that all of the following maps are nullhomotopic
\[ ab,\quad de,\quad a  \langle b,{}^{e_2} c,{}^{e_3} d \rangle,\quad \langle b,{}^{e_2} c,{}^{e_3} d \rangle e. \]

Consider the following expression:
\begin{align}
  \langle \langle a, b, c \rangle, d, e \rangle \pm \langle a, \langle b, c, d \rangle, e \rangle \pm \langle a, b, \langle c, d, e \rangle \rangle
\end{align}

\begin{enumerate}
\item[(c)] Confirm that the expression in (1) is defined,
  adding in the omitted suspensions as necessary.
\item[(d)] Determine an orientation of the signs so that (1) contains zero.
\item[(e)] Identify a collection of nullhomotopies so that the expression is exactly zero.
\end{enumerate}


\section*{\bf Background}

The first $13$ stable homotopy groups of the $2$-local sphere $\Ss_{(2)}$ can be computed using only the Serre spectral sequence. 
These groups (and the ring structure) are given in the table below:

\begin{center}
  \begin{tabular}{|c|c|c|c|}\hline
    stem & group & generator & relations \\\hline\hline
    0 & $\Z$ & $1$ & \\\hline
    1 & $\Z/2$ & $\eta$ & \\\hline
    2 & $\Z/2$ & $\eta^2$ & \\\hline
    3 & $\Z/8$ & $\nu$ & $\eta^3 = 4\nu$ \\\hline
    4 & $0$ & & \\\hline
    5 & $0$ & & \\\hline
    6 & $\Z/2$ & $\nu^2$ & \\\hline
    7 & $\Z/16$ & $\sigma$ & \\\hline
    8 & $\Z/2 \oplus \Z/2$ & $\eta\sigma$, $\epsilon$ & \\\hline
    9 & $\Z/2 \oplus \Z/2 \oplus \Z/2$ & $\eta^2\sigma$, $\eta\epsilon$, $\mu_9$ & $\nu^3 = \eta^2\sigma + \eta\epsilon$ \\\hline
    10 & $\Z/2$ & $\eta\mu_9$ & \\\hline
    11 & $\Z/8$ & $\zeta$ & $\eta^2\mu_9 = 4\zeta$ \\\hline
    12 & $0$ & & \\\hline
    13 & $0$& & \\\hline
  \end{tabular}
\end{center}

You will want to use this information as you do the next four exercises.

%% {\bf Question 2:} Evaluating Toda brackets \hfill
\section*{\bf Exercise 2: Evaluating Toda brackets}

Toda proved the following formula\footnote{This formula comes from the theory of power operations on $\Ss$ and we will return to this subject at the end of the course.} for Toda brackets of elements in $\pi_*\Ss$:
Given $\alpha, \beta \in \pi_*\Ss$ such that $\alpha\beta = 0$
\begin{itemize}
\item If $|\alpha| = k$ is even (or $k$ is odd and $2\alpha =0$),
  there is an element $Q_1(\alpha) \in \pi_{2k+1}\Ss$ such that
  \[ \langle \alpha, \beta, \alpha \rangle = Q_1(\alpha) \beta + \alpha \blacksquare. \]
  Here $\blacksquare$ is an ideterminate used for keeping track of indeterminacy.
\item If $|\alpha| = k$ is odd, then   
  \[ \langle \alpha, \beta, \alpha \rangle \quad\text{ and }\quad \langle \beta, \alpha, 2\alpha \rangle. \]
  have non-empty intersection.
\end{itemize}
For elements in even degree $Q_1(-)$ is a (not necessarily additive) operation of signature
\[ \pi_k\Ss \to \pi_{2k+1}\Ss. \]
In fact, for $\alpha_1, \alpha_2 \in \pi_k\Ss$ with $k$ even we have
\[ Q_1(\alpha_1 + \alpha_2) = Q_1(\alpha_1) + Q_1(\alpha_2) + \eta \alpha_1 \alpha_2 \]
and for $\alpha_1 \in \pi_{k_1}\Ss$ and $\alpha_2 \in \pi_{k_2}\Ss$ with $k_1,k_2$ even we have
\[ Q_1(\alpha_1 \alpha_2) = \alpha_1^2 Q_1(\alpha_2) + Q_1(\alpha_1) \alpha_2^2 \]

Using the table above, shuffling rules and Toda's formula evaluate the following Toda brackets.
Remember to keep indeterminacy in mind.

\begin{enumerate}
\item[(0)] Compute $Q_1(n)$ for each $n \in \Z \cong \pi_0\Ss$.
\item[(a)] $\langle 2, \eta, 2 \rangle$
\item[(b)] $\langle \eta, 2, \eta \rangle$
\item[(c)] $\langle 2, \eta, \nu \rangle$
\item[(d)] $\langle \eta, \nu, \eta \rangle$
\item[(e)] $\langle \nu, \eta, \nu \rangle$
\item[(f)] $\langle \eta^2, \eta^2, \eta^2 \rangle$
\item[(g)] $\langle \eta, 2, \nu^2 \rangle$
\item[(h)] $\langle \mu_9, 2, \eta \rangle$
\end{enumerate}

\section{\bf Exercise 3: Cells structures}

Like spaces, spectra can be given skeletal filtrations.

\begin{dfn}
  A skeletal filtration on $X$ is a sequence 
  \[ \cdots \to X_{-2} \to X_{-1} \to X_0 \to X_1 \to X_2 \to \cdots \]
  such that
  \begin{enumerate}
  \item $\cof(X_{i-1} \to X_i)$ is a sum of copies of $\Ss^i$,
  \item $0 \cong X_{-\infty} \coloneqq \lim X_i$ and
  \item $X \cong X_\infty \coloneqq \colim X_i$.
  \end{enumerate}
\end{dfn}

Alternatively, we can rotate the cofiber sequences in a skeletal filtration and view
the pushout
\[ \begin{tikzcd}
  \oplus \Ss^{i-1} \ar[r] \ar[d] & X_{i-1} \ar[d] \\
  0 \ar[r] & X_{i}
\end{tikzcd} \]
as specifying a collection of cell attachments that build $X_i$ from $X_{i-1}$.\footnote{Recall that discs are contractible, which is why the bottom left is zero.}


\begin{enumerate}
\item[(a)] Prove that every $2$-local finite spectrum $X$ can be given a skeletal filtration with exactly $\mathrm{dim}_{\F_2}(H_i(X;\F_2))$ many $i$-cells. We call a choice of such a skeletal filtration a minimal skeletal filtration. (Hint: use the Hurewicz theorem)
\item[(b)] Extend the results of part (a) to spectra which are of finite type.\footnote{$X$ is of finite type if it is bounded below and $H_i(X;\Z_{(2)})$ is finitely generated for all $i$.}
\end{enumerate}

The homotopy and homology rings of the spectrum $\F_2$ representing cohomology with $\F_2$ coefficients are given by
\[ \pi_*\F_2 \cong \begin{cases} \F_2 & *=0 \\ 0 & *\neq 0 \end{cases} \qquad\qquad H_*(\F_2; \F_2) \coloneqq \pi_*(\F_2 \otimes \F_2) \cong \F_2[\zeta_1, \zeta_2, \dots] \]
where $|\zeta_i| = 2^{i} - 1$.

\begin{enumerate}
\item[(c)] Determine the attaching maps in the minimal $0$-skeleton of $\F_2$.
\item[(d)] Determine the attaching maps in the minimal $1$-skeleton of $\F_2$.
\item[(e)] Determine the attaching maps in the minimal $2$-skeleton of $\F_2$.
\item[(f)] Determine the attaching maps in the minimal $3$-skeleton of $\F_2$.
\end{enumerate}

The homotopy and homology rings of the spectrum $\Z_{(2)}$ representing cohomology with $\Z_{(2)}$ coefficients are given by
\[ \pi_*\F_2 \cong \begin{cases} \Z_{(2)} & *=0 \\ 0 & *\neq 0 \end{cases} \qquad\qquad H_*(\Z_{(2)}; \F_2) \coloneqq \pi_*(\Z_{(2)} \otimes \F_2) \cong \F_2[\zeta_1^2, \zeta_2, \dots] \]
where $|\zeta_1^2| = 2$ and $|\zeta_i| = 2^{i} - 1$ for $i \geq 2$.
The name of the class $\zeta_1^2$ comes from the fact that the reduction mod $2$ map
\[ \Z_{(2)} \to \F_2 \]
induces the natural injective map of polynomial algebras
\[ \F_2[\zeta_1^2, \zeta_2, \dots] \to \F_2[\zeta_1, \zeta_2, \dots] \]
on $\F_2$-homology.

\begin{enumerate}
\item[(g)] Determine the attaching maps in the minimal $2$-skeleton of $\Z_{(2)}$.
\item[(h)] Determine the attaching maps in the minimal $3$-skeleton of $\Z_{(2)}$.
\end{enumerate}


\section{\bf Exercise 4: Characterizing $\mathrm{KU}$}

$\mathrm{KU}$ is an $\E_\infty$-algebra in $\Sp$ whose associated cohomology theory
classifies stable complex vector bundles.
The ring spectrum $\mathrm{KU}_{(2)}$ satisfies the following properties

\begin{enumerate}
\item $\KU_{(2)}$ is $2$-periodic.
  This means that there is an invertible class $\beta$ in $\pi_2\KU_{(2)}$.
\item The connective cover of $\KU_{(2)}$ is typically called $\ku_{(2)}$.
  The $\E_\infty$-algebra $\ku_{(2)}$ is finite type and 
  its $\F_2$-homology is
  \[ H_*(\ku_{(2)}; \F_2) \coloneqq \pi_*(\F_2 \otimes \ku_{(2)}) \cong \F_2[\zeta_1^{2}, \zeta_2^{2}, \zeta_3, \zeta_4, \dots ] \]
  where
  $|\zeta_1^{2}| = 2$,
  $|\zeta_2^{2}| = 6$
  and $|\zeta_i| = 2^{i} - 1$ for $i \geq 3$.
\item There is a map of $\E_\infty$-algebras $\ku_{(2)} \to \Z_{(2)}$ which induces the natural inclusion
  \[ \F_2[\zeta_1^{2}, \zeta_2^{2}, \zeta_3, \zeta_4, \dots ] \to \F_2[\zeta_1^2, \zeta_2, \dots] \]
  on $\F_2$-homology.
\end{enumerate}


\begin{enumerate}
\item[(a)] Show that the localization of $\ku_{(2)}$ at $\beta$ is $\KU_{(2)}$.
\item[(b)] Determine the cells and attaching maps of a minimal 4 skeleton of $\ku_{(2)}$.
\item[(c)] Using (b), compute $\pi_*\ku_{(2)}$ for $*=0,1,2$.
\item[(d)] Determine the ring $\pi_*\KU_{(2)}$.
\item[(e)] Describe $\beta$ as a map into the $2$-skeleton of $\ku_{(2)}$.
\item[(f)] Conclude that properties (1), (2) and (3) above characterize $\KU_{(2)}$.
\item[(g)] How much can you weaken properties (1), (2) and (3) while still being able to prove (f)?
\end{enumerate}


%% \section{\bf Exercise 5: The Wood cofiber sequence}

%% $\mathrm{KO}$ is an $\E_\infty$-algebra in $\Sp$ whose associated cohomology theory
%% classifies stable vector bundles.
%% Complexification gives a map of $\E_\infty$-algebras
%% \[ \KO \to \KU. \]

%% homotopy groups of both

%% (a) determine what the map does on homotopy.


%% (b) Reduce computing cof eta homotopy to a single bracket.



%% (c)

%% The ring spectrum $\mathrm{KU}$ satisfies the following properties


\bibliographystyle{alpha}
\bibliography{bibliography}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% eval: (visual-line-mode t)
%%% eval: (auto-fill-mode 0)
%%% End:
