\pdfoutput=1\relax
\documentclass{amsart}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\input{Preamble.tex}
\usepackage[margin=1.5in]{geometry}

\newtoggle{draft}
% NB: toggle this to hide comments
%\togglefalse{draft}
\toggletrue{draft}

\iftoggle{draft}{
%\usepackage{showkeys}
\newcommand{\atodo}[1]{\todo[color=blue!40]{#1}}
\newcommand{\ttodo}[1]{\todo[color=green!40]{#1}}
\newcommand{\tcomment}[1]{\todo[inline,color=green!40]{#1}}
\newcommand{\NB}[1]{\todo[color=gray!40]{#1}}
\newcommand{\TODO}[1]{\todo[color=red]{#1}}
}{ % else
\newcommand{\NB}[1]{}
\newcommand{\TODO}[1]{}
\renewcommand{\todo}[1]{}
\renewcommand{\todo}[1]{}
}

\title{Review sheet}

%% \author{Robert Burklund}
%% \address{Department of Mathematical Sciences, University of Copenhagen, Denmark}
%% \email{rb@math.ku.dk}

\begin{document}
\maketitle

The theme of this review sheet is having a general sense of how things work and what operations we can do with them.
Not everything here will be used in the course and much of it won't be used for a while.
The purpose is mostly to record some things that are good to know about.

\begin{enumerate}
\item Category theory
  \begin{enumerate}
  \item adjunctions,
  \item monads and comonads,
  \item enriched categories.    
  \end{enumerate}
\item Behavoir of limits and colimits
  \begin{enumerate}    
  \item universal property for mapping in to a product,
  \item universal property for mapping out of a coproduct,
  \item universal property for mapping in to a pullback,
  \item universal property for mapping out of a pushout,
  \item initial and terminal objects,
  \item universal property for mapping in to a fiber,
  \item universal property for mapping out of a cofiber,
  \item rewriting fibers as cofibers in a stable category
    (in order to use universal properties in the other direction),
  \item cofinality,
  \item filtered diagrams and filtered colimits,
  \item sifted diagram and sifted colimits,
  \item rewriting finite colimits in terms of pushouts,
  \item idempotents and idempotent completion,
  \item absolute colimits.
  \end{enumerate}
\item Finiteness conditions
  \begin{enumerate}
  \item finite spaces,
  \item finite spectra,
  \item finite categories,
  \item atomic,
  \item compact,
  \item dualizable.
  \end{enumerate}
\item Presentability
  \begin{enumerate}
  \item HTT Ch 5,
  \item accessibility,
  \item ind completion,
  \item adjoint functor theorems.
  \end{enumerate}  
\item Symmetric monoidal categories
  \begin{enumerate}
  \item Symmetric monoidal $1$-categories,
  \item Symmetric monoidal categories,
  \item Underlying data of sym mon category (binary tensor product, associator),
  \item Extended powers,
  \item $\E_1$-algebras and $\E_\infty$-algebras,
  \item Underlying data from an $\E_\infty$-algebra (mult. map from extended power),
  \item (left) modules over an $\E_1$-algebra,
  \item monoidal structure on $\Mod(R)$ when $R$ is an $\E_2$-algebra,
  \item sym mon structure on $\Mod(R)$ when $R$ is an $\E_\infty$-algebra.
  \end{enumerate}
\end{enumerate}

\bibliographystyle{alpha}
\bibliography{bibliography}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% eval: (visual-line-mode t)
%%% eval: (auto-fill-mode 0)
%%% End:
